---
title: "One Way ANOVA Random Effects Template"
author: "Professor Wendy Martin"
date: "1/10/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE)
require(lolcat)          
require(car)
require(dplyr)
require(flextable)
options(scipen=99,digits = 10)
options(show.signif.stars=FALSE)   # Turn off * to indicate significance
nqtr <- function(x,d){noquote(t(round.object(x, d)))}
ro <- round.object

blanks <- rep(c(' ', '\n'),2) # This will print two blank lines between plots. 

AddBreak <- function() {
for (i in blanks)
  cat(i)
}
```

## One Way Analysis of Variance - Random Effects

### Experiment Title: Gear Experiment

### Step 1: Import File, Define Factors/Variables
```{r import file, echo=FALSE}
Gear <- read.delim("G:/My Drive/EMEN 5042/Lecture Powerpoint and Pdf/Lecture 9 - One Way ANOVA/Gear.dat")

group.labels<-c("Poly1","Poly2","Poly3","Poly4","Poly5")

Gear$material<-factor(Gear$material, labels = group.labels)

# Name of Data Frame
df<-Gear

# Criterion Measure
cm<-Gear$ctf

# Name of Criterion Measure
ncm<-"Cycles to Failure"

# Factor
factor<-Gear$material

# Name of Factor
nf<-"Polymer Material"

# Number of levels
gn<-5

# Flextable Fit to Word
 i = 16.5 # width of the side borders in the word_document output (in centimeters)
 w = i*0.3937 # width of the side borders in the word_document output (in inches)
 
```

### Step 2: Review Descriptive Summary
### Step 3: Review Normality Tests
### Step 4: Review Box and Whisker Plots

```{r Descriptives, echo=FALSE, fig.width=6}
so <- ro(summary.continuous(cm ~ factor, df,stat.miss=F),4)

boxplot(cm ~ factor,data=df, xlab=nf, ylab=ncm, main=paste(ncm,"by",nf), col="red")
```

```{r Descriptives_Table, echo=FALSE}
so %>%
  flextable() %>%
  add_header_lines(values = "Descriptive Summary") %>%
  width(.,width = (w/(ncol(so)))) %>%
  theme_box()

AddBreak()
```

```{r means_normality_test, echo=FALSE}
# Create output object of the cell means
means<-so$mean

# Test for normality of cell means
someans<-ro(summary.continuous(means),4)
someans$dv.name<-"Means"

someans %>%
  flextable() %>%
  add_header_lines(values = "Descriptive Summary of Means") %>%
  width(.,width = (w/(ncol(someans)))) %>%
  theme_box()
```

**Conclusions regarding normality:**
Within group/level normality exists, as well as the normality of the means.

### Step 5: Conduct One Way Random Effects ANOVA 
```{r Random_Effects_ANOVA, echo=FALSE}
raov.out<-aov(formula = cm ~ factor, data = df)
sourcetable.aov <- summary(raov.out)

# Add Totals to ANOVA Table
anova.temp<-sourcetable.aov[[1]]
totaldf<-sum(anova.temp$Df)
totalss<-sum(anova.temp$`Sum Sq`)
total<-data.frame(totaldf, totalss,"","","")
names(total)<-c("Df","Sum Sq","Mean Sq","F value","Pr(>F)")
anova.out<-rbind(anova.temp,as.numeric(total))
rownames(anova.out)<-paste(c(nf,"Residuals","Total"))
anova.out$Source<-rownames(anova.out)
anova.out<-anova.out[c(6,1,2,3,4,5)]

# Random Effects Analysis
# Calculate Treatment Variance
subgroup.n<-(anova.out$Df[3]+1)/(anova.out$Df[1]+1)
Treat.var<-(anova.out$`Mean Sq`[1]-anova.out$`Mean Sq`[2])/subgroup.n

# Within Variance
MSw<-anova.out$`Mean Sq`[2]

# Total Variance
Total.var<-Treat.var + MSw

# Intraclass Correlation Coefficient
rho.i<-ro(Treat.var/Total.var,4)

# Print Source Table
anova.out %>%
  flextable() %>%
  add_header_lines(values = paste("ANOVA Source Table",ncm,"by",nf)) %>%
  colformat_double(j=c('Sum Sq','Mean Sq', 'F value', 'Pr(>F)'), digits=4, na_str = '') %>%
  add_footer_lines(values = paste("Intraclass Correlation Rho =",rho.i)) %>%
   width(.,width = (w/(ncol(anova.out)))) %>%
   theme_box()
```