---
title: "First Assignment"
author: "cccc"
date: "10/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Please complete all the tasks listed below. After completing the tasks download the .Rmd file and upload in the peer review item for grading.

# Additionally please write text between the code chunks explaining what each code chunk is about.

1.a bulleted list with sub-bullets under at least one item (10 pt)

- bulleted list
  - sub bullets
  - sub bullets
- bullet
- bullet


2.a numbered list (10 pt)

1. First item
2. Second item
3. Third item


3.some words italicized (5 pt)

*italicized* word, then not, then again *italicized*

4.some words in bold print (5 pt)

**bold** not bold **bold**

5.two levels of headers (10 pt)

# This is a first-level header

## This is a second-level header

6.the results of some R code without showing the code (10 pt)

```{r, echo = FALSE}
set.seed(123)
x <- 10
y <- 14
z <- x + y
z
```


7.a table of text at least 3 columns wide and 3 rows long (10 pt)

| Column 1 | Column 2 | Column 3 |
|----------|----------|----------|
| Row 1, Column 1 | Row 1, Column 2 | Row 1, Column 3 |
| Row 2, Column 1 | Row 2, Column 2 | Row 2, Column 3 |
| Row 3, Column 1 | Row 3, Column 2 | Row 3, Column 3 |


8.a plot (show the code that generates the plot) (10 pt)

```{r}
library(ggplot2)

ggplot(mtcars, aes(x = wt, y = mpg)) + 
     geom_point() +
     labs(title = "Scatter plot of car weight and miles per gallon",
          x = "Weight (1000 lbs)",
          y = "Miles per gallon")
```

9.your session info (as output from R) (10 pt)

R version 4.2.1 (2022-06-23)
Platform: x86_64-apple-darwin17.0 (64-bit)
Running under: macOS Ventura 13.3.1

Matrix products: default
LAPACK: /Library/Frameworks/R.framework/Versions/4.2/Resources/lib/libRlapack.dylib

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] stats     graphics  grDevices utils    
[5] datasets  methods   base     

other attached packages:
[1] ggplot2_3.4.2

loaded via a namespace (and not attached):
 [1] rstudioapi_0.14  knitr_1.40      
 [3] magrittr_2.0.3   tidyselect_1.2.0
 [5] munsell_0.5.0    colorspace_2.0-3
 [7] R6_2.5.1         rlang_1.1.0     
 [9] fastmap_1.1.0    fansi_1.0.4     
[11] dplyr_1.1.1      tools_4.2.1     
[13] grid_4.2.1       gtable_0.3.1    
[15] xfun_0.34        utf8_1.2.3      
[17] cli_3.6.0        withr_2.5.0     
[19] htmltools_0.5.3  yaml_2.3.6      
[21] digest_0.6.30    tibble_3.2.1    
[23] lifecycle_1.0.3  farver_2.1.1    
[25] vctrs_0.6.1      glue_1.6.2      
[27] evaluate_0.17    rmarkdown_2.17  
[29] labeling_0.4.2   compiler_4.2.1  
[31] pillar_1.8.1     generics_0.1.3  
[33] scales_1.2.1     pkgconfig_2.0.3 

10.Execution of .Rmd file (10 pt)

cmd + shift + k

11. Upload and execution of knitted pdf (10 pt)

