---
title: "Tidying Data Assignment"
author: ''
date: ''
---
# Please complete all the tasks listed below. After completing the tasks download the .Rmd file and upload in the peer review item for grading.

# Additionally please write text between the code chunks explaining what each code chunk is about.

# Refer the linked online textbook in case of any issues.

Load the tidyverse library
```{r}
library(tidyverse)
```

##### Question 1. 

The built in billboard dataset is not tidy. Describe why it is not tidy and then tidy the dataset.

```{r question-1-response}

# First gather up all the week entries into a row for each week for each song (where there is an entry)
billboard2 <- billboard %>% 
  pivot_longer(
    wk1:wk76, 
    names_to = "week", 
    values_to = "rank", 
    values_drop_na = TRUE
  )
billboard2



# Then, convert the week variable to a number and figure out the date corresponding to each week on the chart
billboard3 <- billboard2 %>%
  mutate(
    week = as.integer(gsub("wk", "", week)),
    date = as.Date(date.entered) + 7 * (week - 1),
    date.entered = NULL
  )
billboard3


# Sort the data by artist, track and week. Here are what your first entries should be (formatting can be different):
#>  A tibble: 5,307 x 5
#   artist  track                   date.entered  week  rank   date
 #    <chr>   <chr>                   <date>       <int> <dbl>   <date>
 #  1 2 Pac   Baby Don't Cry (Keep... 2000-02-26       1    87 2000-02-26
 #  2 2 Pac   Baby Don't Cry (Keep... 2000-02-26       2    82 2000-03-04
 #  3 2 Pac   Baby Don't Cry (Keep... 2000-02-26       3    72 2000-03-11
 #  4 2 Pac   Baby Don't Cry (Keep... 2000-02-26       4    77 2000-03-18
 #  5 2 Pac   Baby Don't Cry (Keep... 2000-02-26       5    87 2000-03-25
 #  6 2 Pac   Baby Don't Cry (Keep... 2000-02-26       6    94 2000-04-01
 #  7 2 Pac   Baby Don't Cry (Keep... 2000-02-26       7    99 2000-04-08
 #  8 2Ge+her The Hardest Part Of ... 2000-09-02       1    91 2000-09-02
 #  9 2Ge+her The Hardest Part Of ... 2000-09-02       2    87 2000-09-09
 # 10 2Ge+her The Hardest Part Of ... 2000-09-02       3    92 2000-09-16
 # … with 5,297 more rows

# A tibble: 5,307 x 6


billboard3 %>% arrange(artist, track, week)
billboard3

```

##### Question 2. 

Tidy the "fish_encounters" dataset of fish spotting by monitoring stations. Make the NA into 0 using the option "values_fill = list(seen = 0)"


```{r question-2-response}
fish_encounters

fish_encounters[is.na(fish_encounters)] <- 0
fish_encounters
```

##### Question 3. 

Import the flowers1 dataset. Tidy and pivot the data. Hint: use "read_csv2()" to read in the dataset 

```{r question-3-response}
# Read the dataset using read_csv2()
flowers_data <- read_csv2("flowers1.csv")

# Convert the "Variable" column to character
flowers_data$Variable <- as.character(flowers_data$Variable)

# Tidy and pivot the data using pivot_longer()
tidy_data <- flowers_data %>%
  mutate(Time = as.numeric(Time), replication = as.numeric(replication)) %>%
  pivot_longer(cols = c(Time, replication, Variable, Value), names_to = "Measurement", values_to = "Value", values_drop_na = TRUE)
```

##### Question 4.

Import the flowers2 dataset. Tidy the dataset by turning the one column into 3 separate columns

```{r question-4-response}
df <- read_csv("flowers2.csv")
df

df_split <- separate(df, Flowers/Intensity;Time, into = c("Flowers", "Intensity", "Time"), sep = ";")
```

##### Question 5.

In the following dataset, turn the implicit missing values to explicit 

```{r question-5-response}
output <- tibble(
      treatment   = c("a", "b", "a", "c", "b"),
      gender   = factor(c("M", "F", "F", "M", "M"), levels = c("M", "F", "O")),
      return = c(1.5, 0.75,  0.5,  1.8,  NA)
    )
output


df <- output %>%
  complete(treatment, gender, fill = list(return = NA))
df
```


#####Question 6.

Import the weather dataset as weather. Use "pivot_longer()" to put the days all in one column, then use "pivot_wider" to separate tmax and tmin into separate columns. Print the summary of the final resulting dataset 

```{r question-6-response}
df <- read_csv('weather.csv')
df
```

###### Question 7. 

Load the built in "anscombe" data frame and use "pivot_longer()" to separate all the x and y columns and categorize them into 4 sets

```{r question-7-response}
anscombe %>%
pivot_longer(
cols = starts_with("x"),
names_to = ("x")) %>%
pivot_longer(
cols = starts_with("y"),
names_to = ("y"),
values_to = ("results")
)
```
