---
title: "Tidying Data Assignment"
author: ''
date: ''
---
# Please complete all the tasks listed below. After completing the tasks download the .Rmd file and upload in the peer review item for grading.

# Additionally please write text between the code chunks explaining what each code chunk is about.

# Refer the linked online textbook in case of any issues.

Load the tidyverse library
```{r}
library(tidyverse)
```

##### Question 1. 

The built in billboard dataset is not tidy. Describe why it is not tidy and then tidy the dataset.

```{r question-1-response}

billboard %>%
  #pivot_longer is used to make week an observable data, rather than each week being its own
  pivot_longer(cols = starts_with("wk"),
               names_to = "week",
               values_to = "rank") %>%
  #this is done to make week a dbl in the tibble, as well as to calculate the date of the chart
  mutate(week = parse_number(week),
         date = as.Date(date.entered) + (7 * (week - 1))) %>%
  #added to clean the data (lots of NAs originally)
  filter(!is.na(rank))


```

##### Question 2. 

Tidy the "fish_encounters" dataset of fish spotting by monitoring stations. Make the NA into 0 using the option "values_fill = list(seen = 0)"


```{r question-2-response}
fish_encounters %>%
  #pivot_wider is used to make each station into a column
  pivot_wider(names_from = station,
              values_from = seen,
              values_fill = list(seen = 0))
```

##### Question 3. 

Import the flowers1 dataset. Tidy and pivot the data. Hint: use "read_csv2()" to read in the dataset 

```{r question-3-response}
# Read the dataset using read_csv2()
flowers1 <- read_csv2("flowers1.csv")

#The problem that I noticed was that variable had two observations that
#would be better separated. By using pivot_wider, each observation was instead 
#made into a column with the value associated with it 
flowers1 %>%
  pivot_wider(names_from = Variable,
              values_from = Value)
```

##### Question 4.

Import the flowers2 dataset. Tidy the dataset by turning the one column into 3 separate columns

```{r question-4-response}
flowers2 <- read_csv2("flowers2.csv")
#The problem was Flowers/Intensity, so using separte(), these were separated
#by the /, and convert = TRUE was used to make it a int/dbl
flowers2 %>%
  separate(`Flowers/Intensity`, into = c("Flowers", "Intensity"),
           sep = "/",
           convert = TRUE)
```

##### Question 5.

In the following dataset, turn the implicit missing values to explicit 

```{r question-5-response}
output <- tibble(
      treatment   = c("a", "b", "a", "c", "b"),
      gender   = factor(c("M", "F", "F", "M", "M"), levels = c("M", "F", "O")),
      return = c(1.5, 0.75,  0.5,  1.8,  NA)
    )

#by using complete(), We could see what was implicity missing and make them more noticeable,
#such as no Female C treatment
output %>%
  complete(treatment, gender)
```


#####Question 6.

Import the weather dataset as weather. Use "pivot_longer()" to put the days all in one column, then use "pivot_wider" to separate tmax and tmin into separate columns. Print the summary of the final resulting dataset 

```{r question-6-response}
weather <- read_csv("weather.csv")
#While the data was hard to understand, I first started by making the days
#a column with pivot_longer(), then filtered out the NA, followed by
#pivoting the tmin and tmax separately
weather %>%
  pivot_longer(cols = starts_with("d"),
               names_to = "days",
               values_to = "temp") %>%
  filter(!is.na(temp)) %>%
  pivot_wider(names_from = element,
              values_from = temp)
```

###### Question 7. 

Load the built in "anscombe" data frame and use "pivot_longer()" to separate all the x and y columns and categorize them into 4 sets

```{r question-7-response}
anscombe %>%
  #I used pivot_longer and the names pattern and longer names_to argument
  #to separate the x/y from the number value, creating sets numbered 1-4
  pivot_longer(cols = everything(),
             names_to = c(".value", "set"),
             names_pattern = "(.)(.)",
             values_to = "value")
```
