 \documentclass[11pt]{article} 
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage[siunitx]{circuitikz}
\usepackage{tikz}
\usepackage{tkz-berge}
\usetikzlibrary{positioning, automata, backgrounds}
\usepackage[colorinlistoftodos, color=orange!50]{todonotes}
\usepackage{hyperref}
\usepackage[numbers, square]{natbib}
\usepackage{fancybox}
\usepackage{epsfig}
\usepackage{soul}
\usepackage[framemethod=tikz]{mdframed}
\usepackage[shortlabels]{enumitem}
\usepackage[version=4]{mhchem}
\usepackage{multicol}
\usepackage{forest}
\usepackage{mathtools}
\usepackage{comment}
\usepackage{enumitem}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{color}
\usepackage[numbers]{natbib}
\usepackage{subfiles}
\usepackage{tkz-berge}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}


\newtheorem{prop}{Proposition}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{cor}{Corollary}[prop]

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{definition}
\newtheorem{required}{Problem}

\theoremstyle{definition}
\newtheorem{ex}{Example}

\newcommand{\interval}[4]{\draw (#2, #1) -- (#3, #1); % Usage: \interval{height}{start}{end}{label}
\draw (#2, #1-0.11) -- (#2, #1+0.11); % draw left whisker
\draw (#3, #1-0.11) -- (#3, #1+0.11); % draw right whisker
\node[] at (#2-0.25, #1) {#4};
}


\setlength{\marginparwidth}{3.4cm}
%#########################################################

%To use symbols for footnotes
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
%To change footnotes back to numbers uncomment the following line
%\renewcommand*{\thefootnote}{\arabic{footnote}}

% Enable this command to adjust line spacing for inline math equations.
% \everymath{\displaystyle}

% _______ _____ _______ _      ______ 
%|__   __|_   _|__   __| |    |  ____|
%   | |    | |    | |  | |    | |__   
%   | |    | |    | |  | |    |  __|  
%   | |   _| |_   | |  | |____| |____ 
%   |_|  |_____|  |_|  |______|______|
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{
\normalfont \normalsize 
\textsc{CSCI 3104 Fall 2021 \\ 
Instructors: Profs. Chen and Layer} \\
[10pt] 
\rule{\linewidth}{0.5pt} \\[6pt] 
\huge Problem Set 9 \\
\rule{\linewidth}{2pt}  \\[10pt]
}
%\author{Your Name}
\date{}

\begin{document}
\definecolor {processblue}{cmyk}{0.96,0,0,0}
\maketitle


%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%FILL IN YOUR NAME%%%%%%%
%%%%%%%%%%AND STUDENT ID%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent
Due Date \dotfill April 8 \\
Name \dotfill \textbf{Julia Troni} \\
Student ID \dotfill \textbf{109280095} \\
Collaborators \dotfill \textbf{Just myself in my hospital room :( }

\tableofcontents

\section{Instructions}
 \begin{itemize}
	\item The solutions \textbf{must be typed}, using proper mathematical notation. We cannot accept hand-written solutions. \href{http://ece.uprm.edu/~caceros/latex/introduction.pdf}{Here's a short intro to \LaTeX.}
	\item You should submit your work through the \textbf{class Canvas page} only. Please submit one PDF file, compiled using this \LaTeX \ template.
	\item You may not need a full page for your solutions; pagebreaks are there to help Gradescope automatically find where each problem is. Even if you do not attempt every problem, please submit this document with no fewer pages than the blank template (or Gradescope has issues with it).

	\item You are welcome and encouraged to collaborate with your classmates, as well as consult outside resources. You must \textbf{cite your sources in this document.} \textbf{Copying from any source is an Honor Code violation. Furthermore, all submissions must be in your own words and reflect your understanding of the material.} If there is any confusion about this policy, it is your responsibility to clarify before the due date. 

	\item Posting to \textbf{any} service including, but not limited to Chegg, Reddit, StackExchange, etc., for help on an assignment is a violation of the Honor Code.
\end{itemize}


\newpage
\section{Standard 24- Hash Tables}

\begin{required} \label{Hash3}
Hash tables and balanced binary trees can be both be used to implement a dictionary data structure, which supports insertion, deletion, and lookup operations. In balanced binary trees containing $n$ elements, the runtime of all operations is $\Theta(\log n)$.  \\

\noindent For each of the following three scenarios, compare the average-case performance of a dictionary implemented with a hash table (which resolves collisions with chaining using doubly-linked lists) to a dictionary implemented with a balanced binary tree.
	
\begin{enumerate}[label=(\alph*)]
\item A hash table with hash function $h_1(x)=1$ for all keys $x$.

\begin{proof}[Answer]
Thus, $h_1(x)$ assigns all n items to the same bucket and hence does not follow uniform hashing assumption. \\
So, insertion is $O(1)$ assuming we prepend the current element to our linked list at every collision. However, lookup and delete are both $O(n)$ \\
Since a dictionary implemented with a BBT is $O(logn)$ for insertion, lookup, and delete operations, we have it that a BBT will perform more efficiently for lookup and delete, but less efficiently at insertion. 
\end{proof}


\vskip 15pt
\item A hash table with a hash function $h_2$ that satisfies the Simple Uniform Hashing Assumption, and where the number $m$ of buckets is $\Theta(n)$.
\begin{proof}[Answer]
The expected lookup, search and delete behavior for this hash table is found using $O(1+\alpha)$ where $\alpha= n/m$ \\
Thus,$ \alpha= n/m = n/n=1$ \\
So, insertion is $O(1)$ assuming we prepend the current element to our linked list at every collision. Lookup and delete are both $O(1+1)= O(1)$ \\
Since a dictionary implemented with a BBT is $O(logn)$ for insertion, lookup, and delete operations, we have it that this hash table performs more efficiently for insertion, deletion, and lookup compared to a BBT. 

\end{proof}


\vskip 15pt	
\item A hash table with a hash function $h_3$ that satisfies the Simple Uniform Hashing Assumption, and where the number $m$ of buckets is $\Theta(n^{3/4})$.

\begin{proof}[Answer]
The expected lookup, search and delete behavior for this hash table is found using $O(1+\alpha)$ where $\alpha= n/m$ \\
Thus, $\alpha= n/m = n/n^{3/4}=n^{1/4}$ \\
So, insertion is $O(1)$ assuming we prepend the current element to our linked list at every collision. Lookup and delete are both $O(1+n^{1/4})$ \\
Since a dictionary implemented with a BBT is $O(logn)$ for insertion, lookup, and delete operations, we have it that this hash table performs more efficiently for insertion, deletion, and lookup compared to a BBT.


\end{proof}

\end{enumerate}
\end{required}



\newpage
\section{Standard 25- Doubling Lists and Amortized Analysis}

\begin{required} \label{Hash2}
Consider a hash table $A$ that holds data from a fixed, finite universe $U$. If the hash table starts empty with $b=10$ buckets and doubles its number of buckets whenever the load factor exceeds $\frac{1}{2}$, what is the load factor after adding $n = 36$ elements to the hash table?
\end{required}

\begin{proof}[Answer]
\begin{itemize}
\item Observe that for the first 5 elements $k_1,...k_5, \alpha \leq 5/10 = 1/2$. When we add $k_6$, $\alpha=6/10 > 1/2$. So, we double the size of the array and now we have an array of size $m=20$
\item Now when we add elements $k_7,...k_10, \alpha \leq 10/20 = 1/2$. When we add $k_{11}$, $\alpha=11/20 > 1/2$. So, we double the size of the array and now we have an array of size $m=40$
\item Next when we add elements $k_12,...k_20, \alpha \leq 20/40 = 1/2$. When we add $k_{21}$, $\alpha=21/40 > 1/2$. So, we double the size of the array and now we have an array of size $m=80$
\item Likewise when we add elements $k_22,...k_40, \alpha \leq 20/40 = 1/2$. When we add $k_{41}$, $\alpha=41/80 > 1/2$. So, at element 41 we would double the size of the array to make $m=160$.
\end{itemize}
Since element 36 is greater than 21 but less than 41, the size of the array will be $m=80$. \\
Thus, the load factor, $\alpha= n/m = 36/80= 9/20$
\end{proof}

\end{document} % NOTHING AFTER THIS LINE IS PART OF THE DOCUMENT



