---
title: "Final Project - Covid Analysis"
author: "GAE"
date: "2023-03-01"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load Packages
```{r load packages, echo=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
```

## Covid Data Analysis Project
This project will download the United States Covid Death and Infection data from
John's Hopkins University Github page.  The aim of this project to review the data
for the state of Colorado to identify the counties with the highest infection rate.
The question we will seek to answer is whether the infection rate of the

```{r obtain data from JHU, echo=FALSE}
# Read COVID death data from JHU Github page

US_covid_deaths_url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"

US_covid_confirmed_url = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"

# using the readr library to eliminate issues in parsing dates
raw_US_covid_deaths = readr::read_csv(US_covid_deaths_url)
raw_US_covid_confirmed = readr::read_csv(US_covid_confirmed_url)

# reading data using read.csv yields dates with 'X' appended requiring parsing
# utilizing different library (readr) that corrects issue
#US_covid_deaths = read.csv(url)
```
## Insepct the data
The next step it to review the data that was downloaded to understand the format
for cleaning and tidying.

```{r inspect the data}

head(raw_US_covid_deaths)
head(raw_US_covid_confirmed)

```
# Transform the data
In this next section, we clean and transform the death and confirmed cases data
sets by removing unnecessary columns and summing the cumulative deaths.  The date
column was then converted to a datetime object for plotting and analysis purposes.
The column headings were changed to a more descriptive title. Next the two
cleaned data sets were merged to combine the data on the dates.

```{r clean, transform, and organize the data}
# Pivot wide-format data for dates and sum totals for each state
US_covid_deaths = raw_US_covid_deaths %>% 
  pivot_longer(cols = 13:ncol(raw_US_covid_deaths), names_to = "date") %>%
  group_by(Province_State, Admin2, date,Population) %>%
  summarise("cumulative_deaths" = sum(value, na.rm = TRUE))

US_covid_confirmed = raw_US_covid_confirmed %>% 
  pivot_longer(cols = 12:ncol(raw_US_covid_confirmed), names_to = "date") %>%
  group_by(Province_State, Admin2, date) %>%
  summarise("cumulative_confirmed" = sum(value, na.rm = TRUE))

# Convert dates to datetime object
US_covid_deaths$date = lubridate::mdy(US_covid_deaths$date)
US_covid_confirmed$date = lubridate::mdy(US_covid_confirmed$date)

# Rename columns from Province_State -> State & Admin2 -> County
US_covid_deaths = US_covid_deaths %>% 
  rename_at('Province_State', ~'State') %>%
  rename_at('Admin2', ~'County')

US_covid_confirmed = US_covid_confirmed %>% 
  rename_at('Province_State', ~'State') %>%
  rename_at('Admin2', ~'County')

head(US_covid_deaths,3)
head(US_covid_confirmed,3)

merged_data = merge(x=US_covid_deaths, y=US_covid_confirmed, all.x=TRUE)

CO = merged_data[merged_data$State == "Colorado",]

```

#Filter the data
The merged data set was then filtered by the top three counties for deaths per 
capita and raw total deaths by county.
``` {r Filter Data for Analysis}

# Filter US data frame by State
#state = "Colorado"
#CO = US_covid_deaths[US_covid_deaths$State == state,]

head(CO)

#Group by max value
County_Totals = CO %>% 
  group_by(County, Population) %>% 
  summarise("death_totals" = max(cumulative_deaths)) %>%
  #summarise("population" = min(Population)) %>%
  mutate("deaths_per_capita" = death_totals / Population * 100000) %>%
  arrange(desc(death_totals))

County_Totals

# Collect the ten counties with highest Covid deaths for masking
max_ten_counties = County_Totals[1:10,]
#max_ten_counties

#Select the three highest counties by raw death count and deaths per 100,000 ppl
top_three_by_death_totals = max_ten_counties[1:3,]
#top_three_by_death_totals

top_three_per_capita = max_ten_counties %>%
    arrange(desc(deaths_per_capita))

top_three_per_capita = top_three_per_capita[1:3,]
#top_three_per_capita

#Subset data by raw death totals and per capita

top_three_by_death_totals_df = 
  CO[CO$County %in% top_three_by_death_totals$County,]

top_three_per_capita_df = 
  CO[CO$County %in% top_three_per_capita$County,]

```
<br>
## Data Plots
The following data plots illustrate the cumulative death across the counties of
Colorado.
<br>

```{r data timeseries plots, echo=FALSE}
# Plot data by county
#Time Series Cumulative Deaths by County
CO %>% 
  ggplot(aes(x = date, y = cumulative_deaths, col = County)) +
  ggtitle("Cumulative Deaths by County Time Series Data") + 
  geom_line() +
  theme(legend.position = "none")

#Time Series Cumulative Deaths by County - Top 3 by Death Totals
top_three_by_death_totals_df %>% 
  ggplot(aes(x = date, y = cumulative_deaths, col = County)) +
  ggtitle("Cumulative Deaths by County Time Series Data") + 
  geom_line() + 
  scale_y_continuous(trans="log10")

#Time Series Cumulative Deaths by County - Top 3 by Per Capita
top_three_per_capita_df %>% 
  ggplot(aes(x = date, y = cumulative_deaths, col = County)) +
  ggtitle("Cumulative Deaths by County Time Series Data") + 
  geom_line() +
  scale_y_continuous(trans="log10")
```
<br>
## Time Series Plots
The above time-series plots illustrate the growth/trajectory of the cumulative deaths
by Colorado counties throughout the pandemic.  The first plot shows trajectory of all 64
Colorado counties. The second plot shows the top 3 counties by raw cumulative death count.
The third plot shows the top three counties for cumulative death where the cumulative
death data was normalized by population - this plot illustrates the deaths per 100,000 people.
<br>

``` {Bar Plots}

#Bar Chart Cumulative Deaths by County
CO %>% 
  ggplot(aes(x = County, y=cumulative_deaths, col=County)) +
  geom_bar(position = 'dodge2', stat = "identity") +
  ggtitle("Cumulative Deaths by County Bar Chart")
  theme(axis.text.x=element_text(angle=60,hjust=1), legend.position = "none") 

#Bar plot of the counties with the greatest # of covid deaths
max_ten_counties %>%
  ggplot(aes(x = County, y = death_totals, col = County)) +
  geom_bar(position = 'dodge2', stat = "identity") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  ggtitle("Top 10 Counties for Highest Covid Deaths by Count")

#Per capita death totals
max_ten_counties %>%
  ggplot(aes(x = County, y = deaths_per_capita, col = County)) +
  geom_bar(position = 'dodge2', stat = "identity") +
  theme(axis.text.x=element_text(angle=60,hjust=1)) +
  ggtitle("Top 10 Counties Covid Deaths per 100,000 People")

```
<br>
## Bar Plots
The above bar plots illustrate the cumulative deaths by all counties, the top 10 
higest death count by raw cumulative deaths, and the top 10 counties with the higest
cumulative deaths normalized by population.  This data shows that the raw death counts
were the higest in the largest counties but when normalized, the death rate was not
necessarily the greatest of all counties.


``` {r analysis}

days = max(top_three_by_death_totals_df$date) - 
  min(top_three_by_death_totals_df$date)

day_range = seq(0,days)
#Filter data by top counties for deaths for model analysis
ElPaso = top_three_by_death_totals_df[top_three_by_death_totals_df$County == "El Paso",]
Pueblo = top_three_per_capita_df[top_three_per_capita_df$County == "Pueblo",]

#top_three_by_death_totals_df[top_three_by_death_totals_df$County == "El Paso",] %>%
#  arrange(date) %>%
#  ggplot(aes(x = day_range, y = log(cumulative_deaths), col = County)) +
#  ggtitle("Cumulative Deaths by County Time Series Data") + 
#  geom_point()

#top_three_per_capita_df[top_three_per_capita_df$County == "Pueblo",] %>%
#  arrange(date) %>%
#  ggplot(aes(x = day_range, y = log(cumulative_deaths), col = County)) +
#  ggtitle("Cumulative Deaths by County Time Series Data") + 
#  geom_point()

# Develop linear model for El Paso County
ElPaso_Model = lm(cumulative_deaths ~ cumulative_confirmed, data = ElPaso)

summary(ElPaso_Model)

ElPaso$predicted_deaths = predict(ElPaso_Model)
# Plot model data against actual data
ElPaso %>% 
  ggplot(aes(x=cumulative_confirmed, y=cumulative_deaths)) + 
  geom_point(color='red') +
  geom_smooth(method='lm')


Pueblo_Model = lm(cumulative_deaths ~ cumulative_confirmed, data = Pueblo)

summary(Pueblo_Model)

Pueblo$predicted_deaths = predict(Pueblo_Model)

Pueblo %>% 
  ggplot(aes(x=cumulative_confirmed, y=cumulative_deaths)) + 
  geom_point(color='green') +
  geom_smooth(method='lm')


```
<br>
## Analysis and Conclusion
Linear models were developed for two counties, El Paso and Pueblo.  These models
were develope to determine if the death rate was greater in one county versus the other.
El Paso county had the highest raw count of covid deaths. Pueblo county had the highest
normalized count of covid deaths.  Thus, these two counties were selected for analysis
to determine if one had a higher rate of covid deaths.  When looking at the normalized 
deaths - deaths per 100,000 people - the data indicated that Pueblo county had the 
a much higher rate of covid deaths when compared to El Paso.  The summary data for 
the linear models show that the slope of the line of covid deaths to covid confirmed cases
is much higher for Pueblo when compared to El Paso by approximately 2.88 times.
<br>

