---
title: "String Manipulation and Regular Expressions Practice"
author: ''
---

##### Required Libraries

Load the stringr library

```{r libraries}

```

##### Question 1.

How would you write the following expression? Use writeLines() to see the raw result.

```{r question-1-response}

### She said: "Let's all go!" \never mind! I said "I don't want to."



```

##### Question 2.

Use str_trim() and str_squish() to replace the string: 

"   Help me to      get rid    of extra spaces?      "

with the following strings: 

"Help me to      get rid    of extra spaces?      "
"   Help me to      get rid    of extra spaces?"
"Help me to      get rid    of extra spaces?"
"Help me to get rid of extra spaces?"

```{r quetsion-2-response}

```

##### Question 3. 

Using the words provided in stringr::words, create regular expressions that find all words that:

```{r question-3-response}

# Start with "y"


# End with “x”


# End with "ed", but not with "eed"


# Start or end with "x"


# Contains three or more vowels in a row


# Start and end with the same character


# Contain at least 4 of the 5 vowels


```

##### Question 4. 

Using the sentences provided in stringr::sentences, modify the regex so it no longer picks up the word flickered as a color (which was the case in the lecture video).

```{r question-4-response}

# Show all the sentences that contain at least one color with the word highlighted using str_view_all()

# Use str_extract_all() to produce an array with the colors that appear multiple times in a sentence
```

##### Question 5. 

Load the ISLR package. Then take "names <- row.names(College)"

```{r question-5-response}

# Show the first twenty rows that result if you split Names at each space into 3 parts. 

# Do the same by splitting the first 20 rows on every word boundary.

```
