---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.11.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---


# Lab: Deep Learning

In this section, we show how to fit the examples discussed in the text. We use the `luz` package, which interfaces to the
`torch` package which in turn links to efficient
`C++` code in the LibTorch library. 

This version of the lab was produced by Daniel Falbel and Sigrid
Keydana, both data scientists at **Rstudio** where these packages were
produced.

An advantage over our original `keras` implementation is that this
version does not require a separate `python` installation.

## Single Layer Network on Hitters Data

We start by fitting the models in Section 10.6.
We set up the data, and separate out a training and
test set.

```{r}
library(ISLR2)
Gitters <- na.omit(Hitters)
n <- nrow(Gitters)
set.seed(13)
ntest <- trunc(n / 3)
testid <- sample(1:n, ntest)
```

The linear model should be familiar, but we present it anyway.

```{r}
lfit <- lm(Salary ~ ., data = Gitters[-testid, ])
lpred <- predict(lfit, Gitters[testid, ])
with(Gitters[testid, ], mean(abs(lpred - Salary)))
```

Notice the use of the `with()` command: the first argument is a
dataframe, and the second an expression that can refer to elements of
the dataframe by name. In this instance the dataframe corresponds to
the test data and the expression computes the mean absolute prediction
error on this data.

Next we fit the lasso using `glmnet`. Since this package does
not use formulas, we create `x` and `y` first.

```{r}
x <- scale(model.matrix(Salary ~ . - 1, data = Gitters))
y <- Gitters$Salary
```

The first line makes a call
to `model.matrix()`,  which produces the same matrix that was
used by `lm()` (the `-1`  omits the intercept).
This function automatically converts factors to dummy variables.
The `scale()` function standardizes the matrix so each column
has mean zero and variance one.

```{r}
library(glmnet)
cvfit <- cv.glmnet(x[-testid, ], y[-testid],
    type.measure = "mae")
cpred <- predict(cvfit, x[testid, ], s = "lambda.min")
mean(abs(y[testid] - cpred))
```

To fit the neural network, we first set up a model structure
that describes the network.

```{r}
library(torch)
library(luz) # high-level interface for torch
library(torchvision) # for datasets and image transformation
library(torchdatasets) # for datasets we are going to use
library(zeallot)
torch_manual_seed(13)
```

```{r}
modnn <- nn_module(
  initialize = function(input_size) {
    self$hidden <- nn_linear(input_size, 50)
    self$activation <- nn_relu()
    self$dropout <- nn_dropout(0.4)
    self$output <- nn_linear(50, 1)
  },
  forward = function(x) {
    x %>% 
      self$hidden() %>% 
      self$activation() %>% 
      self$dropout() %>% 
      self$output()
  }
)
```

We have created a model called `modnn` by defining the `initialize()` and `forward()` functions and passing them to the `nn_module()` function. The `initialize()` function is responsible for initializing the submodules that are used by the model. In the `forward` method we implement what happens when the model is called on input data. In this case we use the layers we defined in `initialize()` in that specific order.

`self` is a list-like special object that is used to share information between the methods of the `nn_module()`. When you assign an object to `self` in `initialize()`, it can then be accessed by `forward()`.

The `pipe` operator `%>%`
 passes the previous term as the first argument to the next
function, and returns the result.

We illustrate the use of the pipe operator on a simple example. Earlier, we created `x` using the command

```{r}
x <- scale(model.matrix(Salary ~ . - 1, data = Gitters))
```

We first make a matrix, and then we center and scale each of the variables.
Compound expressions like this can be difficult to parse. We could have obtained the same result using the pipe operator:

```{r}
x <- model.matrix(Salary ~ . - 1, data = Gitters) %>% scale()
```

Using the pipe operator makes it easier to follow the sequence of operations.

We now return to our neural network. The object `modnn` has a single hidden layer with 50 hidden units, and
a ReLU activation function. It then has a dropout layer, in which a
random 40% of the 50 activations from the previous layer are set to zero
during each iteration of the stochastic gradient descent
algorithm. Finally, the output layer has just one unit with no
activation function, indicating that the model provides a single
quantitative output.

Next we add details to  `modnn` that control the fitting
algorithm. We minimize squared-error loss as in
(10.22). The algorithm
tracks the mean absolute error on the training data, and
on validation data if it is supplied.

```{r}
modnn <- modnn %>% 
  setup(
    loss = nn_mse_loss(),
    optimizer = optim_rmsprop,
    metrics = list(luz_metric_mae())
  ) %>% 
  set_hparams(input_size = ncol(x))
```

In the previous line, the pipe operator passes `modnn` as the first argument to `setup()`. 
The `setup()` function embeds these specification into a new model object. 
We also use `set_hparam()` to specify the arguments that should be passed to the 
`initialize()` method of `modnn`. 

Now we fit the model. We supply the training data and the number of `epochs`. By default,
at each step of SGD, the algorithm randomly selects 32 training observations for 
the computation of the gradient. Recall from Sections 10.4 and 10.7
that an epoch amounts to the number of SGD steps required to process $n$
observations. Since the training set has
$n=176$, an epoch is $176/32=5.5$ SGD steps. The `fit()` function has an argument
`valid_data`; these data are not used in the fitting,
but can be used to track the progress of the model (in this case reporting
mean absolute error). Here we
actually supply the test data so we can see mean absolute error of both the
training data and test data as the epochs proceed. To see more options
for fitting, use `?fit.luz_module_generator`.

```{r}
fitted <- modnn %>% 
  fit(
    data = list(x[-testid, ], matrix(y[-testid], ncol = 1)),
    valid_data = list(x[testid, ], matrix(y[testid], ncol = 1)),
    epochs = 20
  )
```
*(Here and elsewhere we have reduced the number of epochs to make
    runtimes manageable; users can of course change back)*

We can plot the `fitted` model to display the mean absolute error for the training and test data. 

```{r}
plot(fitted)
```

Finally, we predict from the final model, and
evaluate its performance on the test data. Due to the use of SGD, the results vary slightly with each
fit. 

```{r}
npred <- predict(fitted, x[testid, ])
mean(abs(y[testid] - npred))
```

## Multilayer Network on the MNIST Digit Data

The `torchvision` package comes with a number of example datasets,
including the `MNIST` digit data. Our first step is to load the
`MNIST` data. The `mnist_dataset()` function is provided for this purpose.

This functions returns a `dataset()`, a data structure implemented in `torch`
allowing one to represent any dataset without making assumptions on where the data is stored and how the data is organized. Usually, torch datasets also implement the 
data acquisition process, like downloading and caching some files on disk.

```{r}
train_ds <- mnist_dataset(root = ".", train = TRUE, download = TRUE)
test_ds <- mnist_dataset(root = ".", train = FALSE, download = TRUE)

str(train_ds[1])
str(test_ds[2])

length(train_ds)
length(test_ds)
```

There are 60,000 images in the training data and 10,000 in the test data. The images are $28\times 28$, and stored as matrix of pixels. We need to transform each one into a vector.  

Neural networks are somewhat sensitive to the scale of the inputs. For example, ridge and
lasso regularization are affected by scaling.  Here the inputs are eight-bit
grayscale values between 0 and 255, so we rescale to the unit
interval. (Note: eight bits means $2^8$, which equals 256. Since the convention
is to start at $0$, the possible values  range from $0$ to $255$.)

To apply these transformations we will re-define `train_ds` and `test_ds`, now passing a the `transform` argument that will apply a transformation to each of
the image inputs.

```{r}
transform <- function(x) {
  x %>% 
    torch_tensor() %>% 
    torch_flatten() %>% 
    torch_div(255)
}
train_ds <- mnist_dataset(
  root = ".", 
  train = TRUE, 
  download = TRUE, 
  transform = transform
)
test_ds <- mnist_dataset(
  root = ".", 
  train = FALSE, 
  download = TRUE,
  transform = transform
)
```

Now we are ready to fit our neural network.

```{r}
modelnn <- nn_module(
  initialize = function() {
    self$linear1 <- nn_linear(in_features = 28*28, out_features = 256)
    self$linear2 <- nn_linear(in_features = 256, out_features = 128)
    self$linear3 <- nn_linear(in_features = 128, out_features = 10)
    
    self$drop1 <- nn_dropout(p = 0.4)
    self$drop2 <- nn_dropout(p = 0.3)
    
    self$activation <- nn_relu()
  },
  forward = function(x) {
    x %>% 
      
      self$linear1() %>% 
      self$activation() %>% 
      self$drop1() %>% 
      
      self$linear2() %>% 
      self$activation() %>% 
      self$drop2() %>% 
      
      self$linear3()
  }
)
```

We define the `intialize()` and `forward()` methods of the `nn_module()`.

In `initialize` we specify all layers that are used in the model.
For example, `nn_linear(784, 256)` defines a dense layer that goes from
$28\times28=784$ input units to a hidden layer of $256$ units. The model
will have 3 of them, each one decreasing the number of output units. The last
will have 10 output units, because each unit will be associated to a different
class, and we have a 10-class classification problem.
We also defined dropout layers using `nn_dropout()`. These will be used
to perform dropout regularization. Finally we define the activation
layer using `nn_relu()`.

In `forward()` we define the order in which these layers are called. We call them in blocks like (linear, activation, dropout), except for the last layer that does not
use an activation function or dropout.

Finally, we use `print` to summarize the model, and to make sure we got it
all right.

```{r}
print(modelnn())
```

The parameters for each layer include a bias term, which results in a
parameter count of 235,146. For example, the first hidden
layer involves $(784+1)\times 256=200{,}960$ parameters.

Next, we add details to the model to specify the fitting algorithm. We fit the model by minimizing the cross-entropy function given by (10.13).

Notice that in `torch` the cross entropy function is defined in terms of 
the logits, for numerical stability and memory efficiency reasons. It does not require the target to be one-hot encoded.

```{r}
modelnn <- modelnn %>% 
  setup(
    loss = nn_cross_entropy_loss(),
    optimizer = optim_rmsprop, 
    metrics = list(luz_metric_accuracy())
  )
```

Now we are ready to go. The final step is to supply training data, and fit the model.

```{r}
system.time(
   fitted <- modelnn %>%
      fit(
        data = train_ds, 
        epochs = 5,
        valid_data = 0.2,
        dataloader_options = list(batch_size = 256),
        verbose = FALSE
      )
 )
plot(fitted)
```

We have suppressed the output here. The output is a progress report on the
fitting of the model, grouped by epoch. This is very useful, since on
large datasets fitting can take time. Fitting this model took 215
seconds on a 2.7GHz MacBook Pro with 4 cores and 16 GB of RAM.
Here we specified a
validation split of 20%, so training is actually performed on
80% of the 60,000 observations in the training set. This is an
alternative to actually supplying validation data, like we did in
Section 10.9.1. See
`?fit.luz_module_generator` for all the optional fitting arguments. SGD  uses batches
of 256 observations in computing the gradient, and doing the
arithmetic, we see that an epoch corresponds to 188 gradient steps.
The last `plot()` command produces a figure similar to Figure 10.18.

To obtain the test error in Table 10.1, we first write
a simple function `accuracy()` that compares predicted and true
class labels, and then use it to evaluate our predictions.

```{r}
accuracy <- function(pred, truth) {
   mean(pred == truth) }

# gets the true classes from all observations in test_ds.
truth <- sapply(seq_along(test_ds), function(x) test_ds[x][[2]])

fitted %>% 
  predict(test_ds) %>% 
  torch_argmax(dim = 2) %>%  # the predicted class is the one with higher 'logit'.
  as_array() %>% # we convert to an R object
  accuracy(truth)
```

The table also reports LDA (Chapter 4) and multiclass logistic
regression. Although packages such as `glmnet` can handle
multiclass logistic regression, they are quite slow on this large
dataset. It is much faster and quite easy to fit such a model
using the `luz` software. We just have an input layer and output layer, and omit the hidden layers!

```{r}
modellr <- nn_module(
  initialize = function() {
    self$linear <- nn_linear(784, 10)
  },
  forward = function(x) {
    self$linear(x)
  }
)
print(modellr())
```

We fit the model just as before.

```{r}
fit_modellr <- modellr %>% 
  setup(
    loss = nn_cross_entropy_loss(),
    optimizer = optim_rmsprop,
    metrics = list(luz_metric_accuracy())
  ) %>% 
  fit(
    data = train_ds, 
    epochs = 5,
    valid_data = 0.2,
    dataloader_options = list(batch_size = 128)
  )

fit_modellr %>% 
  predict(test_ds) %>% 
  torch_argmax(dim = 2) %>%  # the predicted class is the one with higher 'logit'.
  as_array() %>% # we convert to an R object
  accuracy(truth)


# alternatively one can use the `evaluate` function to get the results
# on the test_ds
evaluate(fit_modellr, test_ds)
```

### Convolutional Neural Networks

In this section we fit a CNN to the `CIFAR` data, which is available in the `torchvision`
package. It is arranged in a similar fashion as the `MNIST` data.

```{r}
transform <- function(x) {
  transform_to_tensor(x)
}

train_ds <- cifar100_dataset(
  root = "./", 
  train = TRUE, 
  download = TRUE, 
  transform = transform
)

test_ds <- cifar100_dataset(
  root = "./", 
  train = FALSE, 
  transform = transform
)

str(train_ds[1])
length(train_ds)
```

The CIFAR dataset consists of 50,000 training images, each represented by a 3d tensor:
each three-color image is represented as a set of three channels, each of which consists of
$32\times 32$ eight-bit pixels. We standardize as we did for the
digits, but keep the array structure. This is accomplished with the `transform` argument.

Before we start, we look at some of the training images; similar code produced
Figure 10.5 on page 411.

```{r}
par(mar = c(0, 0, 0, 0), mfrow = c(5, 5))
index <- sample(seq(50000), 25)
for (i in index) plot(as.raster(as.array(train_ds[i][[1]]$permute(c(2,3,1)))))
```

The `as.raster()` function converts the feature map so that it can be plotted as a color image.

Here we specify a moderately-sized  CNN for
demonstration purposes, similar in structure to Figure 10.8.

```{r}
conv_block <- nn_module(
  initialize = function(in_channels, out_channels) {
    self$conv <- nn_conv2d(
      in_channels = in_channels, 
      out_channels = out_channels, 
      kernel_size = c(3,3), 
      padding = "same"
    )
    self$relu <- nn_relu()
    self$pool <- nn_max_pool2d(kernel_size = c(2,2))
  },
  forward = function(x) {
    x %>% 
      self$conv() %>% 
      self$relu() %>% 
      self$pool()
  }
)

model <- nn_module(
  initialize = function() {
    self$conv <- nn_sequential(
      conv_block(3, 32),
      conv_block(32, 64),
      conv_block(64, 128),
      conv_block(128, 256)
    )
    self$output <- nn_sequential(
      nn_dropout(0.5),
      nn_linear(2*2*256, 512),
      nn_relu(),
      nn_linear(512, 100)
    )
  },
  forward = function(x) {
    x %>% 
      self$conv() %>% 
      torch_flatten(start_dim = 2) %>% 
      self$output()
  }
)
model()
```

Notice that we used the `padding = "same"` argument to
`nn_conv2d()`, which ensures that the output channels have the
same dimension as the input channels. There are 32 channels in the first
hidden layer, in contrast to the three channels in the input layer. We
use a $3\times 3$ convolution filter for each channel in all the layers. Each
convolution is followed by a max-pooling layer over $2\times2$
blocks. By studying the summary, we can see that the channels halve in both
dimensions
after each of these max-pooling operations. After the last of these we
have a layer with  256 channels of dimension $2\times 2$. These are then
flattened to a dense layer of size 1,024:
in other words, each of the $2\times 2$ matrices is turned into a
$4$-vector, and put side-by-side in one layer. This is followed by a
dropout regularization layer,  then
another dense layer of size 512, and finally, the
output layer.

Finally, we specify the fitting algorithm, and fit the model.

```{r}
fitted <- model %>% 
  setup(
    loss = nn_cross_entropy_loss(),
    optimizer = optim_rmsprop, 
    metrics = list(luz_metric_accuracy())
  ) %>% 
  set_opt_hparams(lr = 0.001) %>% 
  fit(
    train_ds,
    epochs = 10, #30,
    valid_data = 0.2,
    dataloader_options = list(batch_size = 128)
  )

print(fitted)

evaluate(fitted, test_ds)
```

This model takes 10 minutes to run and achieves 36% accuracy on the test
data. Although this is not terrible for 100-class data (a random
classifier gets 1% accuracy), searching the web we see results around
75%. Typically it takes a lot of architecture carpentry,
fiddling with regularization, and time to achieve such results.

## Using Pretrained CNN Models

We now show how to use a CNN pretrained on the  `imagenet` database to classify natural
images, and demonstrate how we produced Figure 10.10.
We copied six jpeg images from a digital photo album into the
directory `book_images`. (These images are available
  from the data section of  [www.statlearning.com](www.statlearning.com), the ISLR book website. Download `book_images.zip`; when
clicked it creates the `book_images` directory.) We first read in the images, and
convert them into the array format expected by the `torch`
software to match the specifications in `imagenet`. Make sure that your working directory in `R` is set to the folder in which the images are stored.

```{r}
img_dir <- "book_images"
image_names <- list.files(img_dir)
num_images <- length(image_names)
x <- torch_empty(num_images, 3, 224, 224)
for (i in 1:num_images) {
   img_path <- file.path(img_dir, image_names[i])
   img <- img_path %>% 
     base_loader() %>% 
     transform_to_tensor() %>% 
     transform_resize(c(224, 224)) %>% 
     # normalize with imagenet mean and stds.
     transform_normalize(
       mean = c(0.485, 0.456, 0.406),
       std = c(0.229, 0.224, 0.225)
     )
   x[i,,, ] <- img
}
```

We then load the trained network. The model has 18 layers, with a fair bit of complexity.

```{r}
model <- torchvision::model_resnet18(pretrained = TRUE)
model$eval() # put the model in evaluation mode
```

Finally, we classify our six images, and return the top three class
choices in terms of predicted probability for each.

```{r}
preds <- model(x)

mapping <- jsonlite::read_json("https://s3.amazonaws.com/deep-learning-models/image-models/imagenet_class_index.json") %>% 
  sapply(function(x) x[[2]])

top3 <- torch_topk(preds, dim = 2, k = 3)

top3_prob <- top3[[1]] %>% 
  nnf_softmax(dim = 2) %>% 
  torch_unbind() %>% 
  lapply(as.numeric)

top3_class <- top3[[2]] %>% 
  torch_unbind() %>% 
  lapply(function(x) mapping[as.integer(x)])

result <- purrr::map2(top3_prob, top3_class, function(pr, cl) {
  names(pr) <- cl
  pr
})
names(result) <- image_names
print(result)
```

## IMDb Document Classification

Now we perform document classification (Section 10.4) on the `IMDB` dataset, which is available as part of the `torchdatasets`
package. We  limit the dictionary size to the
10,000  most frequently-used words and tokens.

```{r}
max_features <- 10000
imdb_train <- imdb_dataset(
  root = ".", 
  download = TRUE, 
  num_words = max_features
)
imdb_test <- imdb_dataset(
  root = ".", 
  download = TRUE, 
  num_words = max_features
)
```

Each element of `imdb_train` is a vector of numbers between 1 and
10000 (the document), referring to the words found in the dictionary.
For example, the first training document is the positive review on
page  419. The indices of the first 12 words are given below.


```{r}
imdb_train[1]$x[1:12]
```

To see the words, we create a function, `decode_review()`, that provides a simple interface to the dictionary.

```{r}
word_index <- imdb_train$vocabulary
decode_review <- function(text, word_index) {
   word <- names(word_index)
   idx <- unlist(word_index, use.names = FALSE)
   word <- c("<PAD>", "<START>", "<UNK>", word)
   words <- word[text]
   paste(words, collapse = " ")
}
decode_review(imdb_train[1]$x[1:12], word_index)
```

Next we write a function to *one-hot* encode each document in a list
of documents, and return a binary matrix in sparse-matrix format.

```{r}
library(Matrix)
one_hot <- function(sequences, dimension) {
   seqlen <- sapply(sequences, length)
   n <- length(seqlen)
   rowind <- rep(1:n, seqlen)
   colind <- unlist(sequences)
   sparseMatrix(i = rowind, j = colind,
      dims = c(n, dimension))
}
```

To construct the sparse matrix, one supplies just the entries that are
nonzero. In the last line we call the function `sparseMatrix()` and supply the
row indices corresponding to each document and the column indices
corresponding to the words in each document, since we omit the
values they are taken to be all ones.
Words that appear more than once in any given document still get
recorded as a one.

```{r}
# collect all values into a list
train <- seq_along(imdb_train) %>% 
  lapply(function(i) imdb_train[i]) %>% 
  purrr::transpose()
test <- seq_along(imdb_test) %>% 
  lapply(function(i) imdb_test[i]) %>% 
  purrr::transpose()

# num_words + padding + start + oov token = 10000 + 3
x_train_1h <- one_hot(train$x, 10000 + 3)
x_test_1h <- one_hot(test$x, 10000 + 3)
dim(x_train_1h)
nnzero(x_train_1h) / (25000 * (10000 + 3))
```

Only 1.3% of the entries are nonzero, so this amounts to considerable
savings in memory.
We create a validation set of size 2,000, leaving 23,000 for training.

```{r}
set.seed(3)
ival <- sample(seq(along = train$y), 2000)
itrain <- seq_along(train$y)[-ival]
```

First we fit a lasso logistic regression model using `glmnet()`
on the training data, and evaluate its performance on the validation
data. Finally, we plot the accuracy, `acclmv`, as a function of
the shrinkage parameter, $\lambda$. Similar expressions compute the
performance on the test data, and were used to produce the left plot
in Figure 10.11.

The code  takes advantage of the sparse-matrix format of
`x_train_1h`, and runs in about 5 seconds; in the usual
dense format it would take about 5 minutes.

```{r}
library(glmnet)
y_train <- unlist(train$y)

fitlm <- glmnet(x_train_1h[itrain, ], unlist(y_train[itrain]),
    family = "binomial", standardize = FALSE)
classlmv <- predict(fitlm, x_train_1h[ival, ]) > 0
acclmv <- apply(classlmv, 2, accuracy,  unlist(y_train[ival]) > 0)
```

We applied the `accuracy()` function that we wrote in Lab 10.9.2
to every column of the prediction matrix
`classlmv`, and since this is a logical matrix of  `TRUE/FALSE` values, we
supply the second argument `truth` as a logical vector as well.

Before making a plot, we adjust the plotting window.

```{r}
par(mar = c(4, 4, 4, 4), mfrow = c(1, 1))
plot(-log(fitlm$lambda), acclmv)
```

Next we fit a fully-connected neural network with two hidden layers,
each with 16 units and ReLU activation.

```{r}
model <- nn_module(
  initialize = function(input_size = 10000 + 3) {
    self$dense1 <- nn_linear(input_size, 16)
    self$relu <- nn_relu()
    self$dense2 <- nn_linear(16, 16)
    self$output <- nn_linear(16, 1)
  },
  forward = function(x) {
    x %>% 
      self$dense1() %>% 
      self$relu() %>% 
      self$dense2() %>% 
      self$relu() %>% 
      self$output() %>% 
      torch_flatten(start_dim = 1)
  }
)
model <- model %>% 
  setup(
    loss = nn_bce_with_logits_loss(),
    optimizer = optim_rmsprop,
    metrics = list(luz_metric_binary_accuracy_with_logits())
  ) %>% 
  set_opt_hparams(lr = 0.001)

fitted <- model %>% 
  fit(
    # we transform the training and validation data into torch tensors
    list(
      torch_tensor(as.matrix(x_train_1h[itrain,]), dtype = torch_float()), 
      torch_tensor(unlist(train$y[itrain]))
    ),
    valid_data = list(
      torch_tensor(as.matrix(x_train_1h[ival, ]), dtype = torch_float()), 
      torch_tensor(unlist(train$y[ival]))
    ),
    dataloader_options = list(batch_size = 512),
    epochs = 10
  )

plot(fitted)  
```
    
The `fitted` object has a `get_metrics` method that
gets both the training and validation accuracy at each epoch.
Figure 10.11 includes test accuracy at each epoch as well. To
compute the test accuracy, we
rerun the entire sequence above, replacing the last line with

```{r}
fitted <- model %>% 
  fit(
    list(
      torch_tensor(as.matrix(x_train_1h[itrain,]), dtype = torch_float()), 
      torch_tensor(unlist(train$y[itrain]))
    ),
    valid_data = list(
      torch_tensor(as.matrix(x_test_1h), dtype = torch_float()), 
      torch_tensor(unlist(test$y))
    ),
    dataloader_options = list(batch_size = 512),
    epochs = 10
  )
```

## Recurrent Neural Networks

In this lab we fit the models illustrated in
Section 10.5.

### Sequential Models for Document Classification


Here we  fit a simple  LSTM RNN for sentiment analysis with
the `IMDB` movie-review data, as discussed in Section 10.5.1. We showed how to input the data in
10.9.5, so we will not repeat that here.

We first calculate the lengths of the documents.


```{r}
wc <- sapply(seq_along(imdb_train), function(i) length(imdb_train[i]$x))
median(wc)
sum(wc <= 500) / length(wc)
```

We see that over 91% of the documents have fewer than 500 words. Our
RNN requires all the document sequences to have the same length. We hence
restrict the document lengths to the last $L=500$ words, and pad the
beginning of the
shorter ones with blanks. We will use `torchdatasets` functionality for this.

```{r}
maxlen <- 500
num_words <- 10000
imdb_train <- imdb_dataset(root = ".", split = "train", num_words = num_words,
                           maxlen = maxlen)
imdb_test <- imdb_dataset(root = ".", split = "test", num_words = num_words,
                           maxlen = maxlen)

vocab <- c(rep(NA, imdb_train$index_from - 1), imdb_train$get_vocabulary())
tail(names(vocab)[imdb_train[1]$x])
```


The last expression shows the last few words in the first document. At this stage, each of the 500 words in the document  is represented using an integer
corresponding to the location of that word in the 10,000-word dictionary.
The first layer of the RNN is an embedding layer of size 32, which will be
learned during  training. This layer one-hot encodes  each document
as a matrix of dimension $500 \times 10,000$, and then maps these
$10,000$ dimensions down to $32$.

```{r}
model <- nn_module(
  initialize = function() {
    self$embedding <- nn_embedding(10000 + 3, 32)
    self$lstm <- nn_lstm(input_size = 32, hidden_size = 32, batch_first = TRUE)
    self$dense <- nn_linear(32, 1)
  },
  forward = function(x) {
    c(output, c(hn, cn)) %<-% (x %>% 
      self$embedding() %>% 
      self$lstm())
    output[,-1,] %>%  # get the last output
      self$dense() %>% 
      torch_flatten(start_dim = 1)
  }
)
```

The second  layer is an LSTM with 32 units, and the output
layer is a single logit for the binary classification task.
The rest is now similar to other networks we have fit. We
track the test performance as the network is fit, and see that it attains 87\% accuracy.

```{r}
model <- model %>% 
  setup(
    loss = nn_bce_with_logits_loss(),
    optimizer = optim_rmsprop,
    metrics = list(luz_metric_binary_accuracy_with_logits())
  ) %>% 
  set_opt_hparams(lr = 0.001)

fitted <- model %>% fit(
  imdb_train, 
  epochs = 10,
  dataloader_options = list(batch_size = 128),
  valid_data = imdb_test
)
plot(fitted)
predy <- torch_sigmoid(predict(fitted, imdb_test)) > 0.5
evaluate(fitted, imdb_test, dataloader_options = list(batch_size = 512))
```

###  Time Series Prediction

We now show how to fit the models in Section 10.5.2
for  time series prediction.
We first set up the data, and standardize each of the variables.

```{r}
library(ISLR2)
xdata <- data.matrix(
 NYSE[, c("DJ_return", "log_volume","log_volatility")]
 )
istrain <- NYSE[, "train"]
xdata <- scale(xdata)
```

The  variable `istrain` contains a `TRUE` for each year that  is in the training set, and a `FALSE` for each year
 in the test set.

We first write functions to create lagged versions of the three time series.  We start with a function that takes as input a data
matrix and a lag $L$, and returns a lagged version of the matrix. It
simply inserts $L$ rows of `NA` at the top, and truncates the
bottom.

```{r}
lagm <- function(x, k = 1) {
   n <- nrow(x)
   pad <- matrix(NA, k, ncol(x))
   rbind(pad, x[1:(n - k), ])
}
```

We now use this function to create a data frame with all the required
lags, as well as the response variable.

```{r}
arframe <- data.frame(log_volume = xdata[, "log_volume"],
   L1 = lagm(xdata, 1), L2 = lagm(xdata, 2),
   L3 = lagm(xdata, 3), L4 = lagm(xdata, 4),
   L5 = lagm(xdata, 5)
 )
```


If we look at the first five rows of this frame, we will see some
missing values in the lagged variables (due to the construction above). We remove these rows, and adjust `istrain`
accordingly.

```{r}
arframe <- arframe[-(1:5), ]
istrain <- istrain[-(1:5)]
```

We now fit the linear AR model to the training data using `lm()`, and predict on the
test data.

```{r}
arfit <- lm(log_volume ~ ., data = arframe[istrain, ])
arpred <- predict(arfit, arframe[!istrain, ])
V0 <- var(arframe[!istrain, "log_volume"])
1 - mean((arpred - arframe[!istrain, "log_volume"])^2) / V0
```

The last two lines compute the $R^2$ on the test data, as defined in (3.17).

We refit this model, including the factor variable `day_of_week`.

```{r}
arframed <-
    data.frame(day = NYSE[-(1:5), "day_of_week"], arframe)
arfitd <- lm(log_volume ~ ., data = arframed[istrain, ])
arpredd <- predict(arfitd, arframed[!istrain, ])
1 - mean((arpredd - arframe[!istrain, "log_volume"])^2) / V0
```

To fit the RNN, we need to reshape these data, since it expects a
sequence of $L=5$ feature vectors $X=\{X_\ell\}_1^L$ for each observation, as in (10.20) on
page  428. These are  lagged versions of the
time series going back $L$ time points.

```{r}
n <- nrow(arframe)
xrnn <- data.matrix(arframe[, -1])
xrnn <- array(xrnn, c(n, 3, 5))
xrnn <- xrnn[,, 5:1]
xrnn <- aperm(xrnn, c(1, 3, 2))
dim(xrnn)
```

We have done this in four steps. The first simply extracts the
$n\times 15$ matrix of lagged versions of the three predictor
variables from `arframe`. The second converts this matrix to a
$n\times 3\times 5$ array. We can do this by simply changing the
dimension attribute, since the new array is filled column wise. The
third step reverses the order of lagged variables, so that index $1$
is furthest back in time, and index $5$ closest. The
final step rearranges the coordinates of the array (like a partial
transpose) into the format that the RNN module in `torch`
expects.

Now we are ready to proceed with the RNN, which uses 12 hidden units.

```{r}
model <- nn_module(
  initialize = function() {
    self$rnn <- nn_rnn(3, 12, batch_first = TRUE)
    self$dense <- nn_linear(12, 1)
    self$dropout <- nn_dropout(0.2)
  },
  forward = function(x) {
    c(output, ...) %<-% (x %>% 
      self$rnn())
    output[,-1,] %>% 
      self$dropout() %>% 
      self$dense() %>% 
      torch_flatten(start_dim = 1)
  }
)

model <- model %>% 
  setup(
    optimizer = optim_rmsprop,
    loss = nn_mse_loss()
  ) %>% 
  set_opt_hparams(lr = 0.001)

```

The output layer has a single unit for the response.

We  fit the model in a similar fashion to previous networks. We
supply the `fit` function with test data as validation data, so that when
we monitor its progress and plot the history function we can see the
progress on the test data. Of course we should not use this as a basis for
early stopping, since then the test performance would be biased.


```{r}
fitted <- model %>% fit(
    list(xrnn[istrain,, ], arframe[istrain, "log_volume"]),
    epochs = 75, #epochs = 200,
    dataloader_options = list(batch_size = 64),
    valid_data =
      list(xrnn[!istrain,, ], arframe[!istrain, "log_volume"])
  )
kpred <- as.numeric(predict(fitted, xrnn[!istrain,, ]))
1 - mean((kpred - arframe[!istrain, "log_volume"])^2) / V0
```

This model takes about one minute to train.

We could replace the  `nn_module()`  command above with the following command:

```{r}
model <- nn_module(
  initialize = function() {
    self$dense <- nn_linear(15, 1)
  },
  forward = function(x) {
    x %>% 
      torch_flatten(start_dim = 2) %>% 
      self$dense()
  }
)
```

Here, `torch_flatten()` simply takes the input sequence and
turns it into a long vector of predictors. This results in a linear AR model.
To fit a nonlinear AR model, we could add in a hidden layer.

However, since we already have the matrix of lagged variables from the AR
model that we fit earlier using the `lm()` command, we can actually fit a nonlinear AR model without needing to perform flattening.
We extract the model matrix `x` from `arframed`, which
includes the `day_of_week` variable.

```{r}
x <- model.matrix(log_volume ~ . - 1, data = arframed)
colnames(x)
```

The `-1` in the formula avoids the creation of a column of ones for
the intercept. The variable `day\_of\_week` is a five-level
factor (there are five trading days), and the
 `-1` results in  five rather than four dummy variables.

The rest of the steps to fit a nonlinear AR model should by now be familiar.

```{r}
arnnd <- nn_module(
  initialize = function() {
    self$dense <- nn_linear(15, 32)
    self$dropout <- nn_dropout(0.5)
    self$activation <- nn_relu()
    self$output <- nn_linear(32, 1)
    
  },
  forward = function(x) {
    x %>% 
      torch_flatten(start_dim = 2) %>% 
      self$dense() %>% 
      self$activation() %>% 
      self$dropout() %>% 
      self$output() %>% 
      torch_flatten(start_dim = 1)
  }
)
arnnd <- arnnd %>% 
  setup(
    optimizer = optim_rmsprop,
    loss = nn_mse_loss()
  ) %>% 
  set_opt_hparams(lr = 0.001)

fitted <- arnnd %>% fit(
    list(xrnn[istrain,, ], arframe[istrain, "log_volume"]),
    epochs = 30, #epochs = 200,
    dataloader_options = list(batch_size = 64),
    valid_data =
      list(xrnn[!istrain,, ], arframe[!istrain, "log_volume"])
  )
plot(fitted)
npred <- as.numeric(predict(fitted, xrnn[!istrain, ,]))
1 - mean((arframe[!istrain, "log_volume"] - npred)^2) / V0
```
